---
title: "Assignment 3 - Task 3"
author: "Jessica West"
date: "10/23/2021"
output: html_document
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(DT)
library(lubridate)
```

```{r}
urchins <- read_csv(here ("data", "urchins.csv"))
urchins_clean <- clean_names(urchins)
```

```{r}
urchins_sum <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(mean_diameter_cm = mean(size, na.rm = TRUE),
            median_diameter_cm = median(size, na.rm = TRUE),
            sd_diameter_cm = sd(size, na.rm = TRUE),
            sample_size = n())

ggplot(data = urchins_clean, aes(x = size)) + 
   facet_wrap(~ common_name) +
  geom_histogram(fill="blue", color = "black", bins = 10) +
  theme_classic() +
   labs(x = "Diameter (cm)", 
       y = "Number of Urchin Samples", 
       title = "Size Distribution of Purple and Red Urchins",
       subtitle = "2008-2020",
       caption = "Data source: Reed, D C (2021)")

urchins_sum %>% 
    kable(digits = 2,
          caption = "Size Distribution of Purple and Red Urchins (2008-2020)", 
          col.names = c("Common Name", "Mean (cm)", "Median (cm)", "Standard Deviation (cm)", "Sample Size")) %>%
  kable_styling(full_width = FALSE)
```


The above figures reflect a not-quite evenly distributed mean for purple urchins and a normal distribution for red urchins diameter size. Purple urchins also show a smaller/narrower standard deviation (~65%) and red urchins show a larger/wider standard deviation (~99%).


```{r}
purple_urchins_only <- urchins_clean %>% 
  filter(common_name == "Purple Urchin")
purple_urchin_dates <- purple_urchins_only %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year_new = year(date_new))
purple_urchins <- purple_urchin_dates %>%
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Isla Vista Reef",
    site == "IVEE" ~ "Mohawk Reef"))
```

```{r}
total_purple_urchins_site_year <- purple_urchins %>% 
  group_by(site_full, year_new) %>% 
  count (year_new)
```

```{r, include = FALSE}
total_purple_urchins_site_year %>% 
  kable(digits = 2) %>%
  kable_styling(full_width = FALSE)
```

```{r}
ggplot(data = total_purple_urchins_site_year, aes(x = year_new , y = n)) + 
   facet_wrap(~ site_full, scales = "free") +
  geom_col(fill="purple", color = "black", bins = 10) +
  theme_classic() +
   labs(x = "Year", 
       y = "Population", 
       title = "Population Trends of Purple Urchin (Strongylocentrotus purpuratus)",
       subtitle = "2008-2020",
       caption = "Data source: Reed, D C (2021)")
```

The figure above indicates what appears to be a decline in the purple urchin population at each location since about 2012. The population visually seems relatively stable after this decline between 2013-2019, but with a significant drop in 2020 (with relatively similar low numbers in 2008, save for Mohwak Reef where samples were not collected this year). While this graph may visually provide insights into urchin numbers, we should also consider external variables that may be influencing urchin populations within each site (ie. kelp density, harvest rates, etc.) in order to make a more accurate assessment of overall population trends.

